Re-run, Repeat, Reproduce, Reuse, Replicate:\\Transforming Code into Scientific Contributions

### 2017-07-25 draft.tex
Fix spelling

Replicability is a cornerstone of science.  If an experimental result cannot be
re-obtained by an independent party, it merely becomes, at best, an observation
that may inspire future research (\cite{Mesirov:2010,osc:2015}). Replication
issues have received increased attention in recent years, with a particular
mediatic focus on medicine and psychology (\cite{Iqbal:2016}).  One could think
that computer science would mostly be shielded from such issues, since a
computer program intrinsically describes precisely what it does and is easy to
transmit perfectly to other researchers.

@@

A \emph{re-runnable} code is one that can be run again when needed, i.e. more than the one time that was needed to produce the results. It is important to notice that the re-runnability of a code is not an intrinsic property. Rather, it depends on the context, and re-runnability becomes increasingly difficult as the code ages. Therefore, to be re-runnable in other researchers' computers and through time, a re-runnable code should describes---with enough details to be recreated---an execution environment in which it is executable. As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.\\

@@

Setting seeds is should be done carefully. Using 439 as a seed in the previous program would result in ten consecutive +1 steps\footnote{With CPython 3.3-3.6. See the next section for details.}, which---although a perfectly valid random walk---lend itself to a gross misinterpretation of the overall dynamic of the algorithm. Verifying that the qualitative aspects of the results and the conclusion we make are not tied to a specific initialization of the pseudo-random generator is an integral part of any scientific undertaking in computational science;
this is usually done by repeating the simulations multiple times with different seeds.
%FIXME: is `simulations` pertinent here?

@@

% scripts to generates figures
The code should also contains all the scripts that were used to produce the figures. While it may be considered extraneous, code for plotting figures may contains non-trivial statistics code, and is actually just as vulnerable to bugs. With a code that is re-runnable, repeatable and contains such scripts, a user can quickly check that published figures can be reproduced.\\

@@

% reproducibility implies availability
It should be obvious by now that \emph{reproducibility implies availability}. As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request. While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable. Code \emph{and} input \emph{and} result files \emph{and} figure data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repository such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment. Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.}.\\

@@

Reusability of course directly benefits others researchers from your team and outside of it. The easier it is to use your code, the lower the threshold is for other to study, modify and extend it. Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage. Moreover, scientists prefer to base their work on results they can verify. A reproducible and reusable code offers a platform both validated and easy-to-use, fostering the development of derivative works by other researchers. Those derivative works contribute to the impact of your original contribution.\\

@@

Despites all this, reusability is often overlooked, and it is not hard to see why. Scientists are rarely trained in software engineering, and reusability can represent an expensive endeavour if undertaken as an afterthought, for little tangible short-term benefits, for a codebase that might, after all, see only a single use. And, in fact, reusability is not as indispensable a requirement as the others one presented here.\\

Yet, some simple measures can tremendously increase reusability, and at the same time strenghten reproducibility and re-runnability.\\

@@

Similarily, code behavior should be changed by commenting/uncommenting code \citep{Wilson:2017}. Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitely set through parameters accessible to the end-user. This improves reproducibility in two ways: it allows those conditions to be explicitely tracked in the result files since they contain the parameters used. It also allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper. With a documention explaining which script or configuration file correspond to which experiment, reproducing the different figures becomes straightforward.\\
%FIXME repeating what was said in reproducible.

Documentation, in fact, is one of the most potent tools for reusability. a proper documentation on how to install and run the software often makes the difference for other reasearchers whether they manage to use the code or not.\\

Furthermore, A comment describing what each function does, however evident, can avoid hours of head-scratching. Great code may need few comments. Scientists, however, are not always brilliant developers. Code quality, in science, averages on mediocre. Of course, bad, complicated code should be rewritten until is simple enough to explain itself. But realistically, this is hardly ever going to be done: there is simply not enough incentive for it. There, a comment that explains the intention and scientific reasons behind a block of code can tremendously useful.\\

@@

Re-usability is not an absolute requirement for scientific code. But it has many benefit, and a few simple measure can foster it considerably.

@@

Here we are be able to replicate exactly the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standart Python library of CPython 3.6.1), or to the code itself.\\

But there are still more sublte problems with the description given above.
If we look more closely at it, we can realize
that nothing is said about the specific case of {\tt 0} when generating a step.
Do whe have to consider {\tt 0} to be a positive or a negative step? Without
further information and without the original code, it is up to the reader to
decide. Likewise, the description is ambiguous regarding the first element of the walk. Is the initialization value included (it was not in our codes so far)? This slight difference changes the statistics, especially of short runs.\\

There is no better way to ferret out the ambiguities, big and small, of an article than to replicate it. This is one of the reason why the \href{https://rescience.github.io}{ReScience
  journal} has been created. This journal {\em targets computational research and
encourages the explicit replication of already published research, promoting
new and open-source implementations in order to ensure that the original
research is reproducible}.\\

@@

%FIXME: can still be improved
Through a small random walk example using Python, we've seen a variety of
pitfalls that are easy to remain unaware of while writing scientific code. The
R$^0$ form, while easy to write and simple to understand is not a good
scientific code. Adding a single comment, we transformed this code into the
R$^1$ version (rerunnable) that partially solved the very common dependencies
problem and will allow us to re-run the program two months or two years from
now. In the meantime, the R$^2$ version (repeatable) guarantee us (to some
extent) we can faithfully repeat our experiment from one run to the other.  Of
course, the full-fledge R$^4$ (reusable) form is much preferred but it requires
a fair amount of work that might be too much of a burden. R$^3$ (reproducible)
form seems to be an acceptable compromise and should be accepted as the minimum
scientific standard \citep{Wilson:2017}. This means this should be actually
checked by reviewers and publishers when code is part of a work worth to be
published. But it's hardly the case today.\\

@@

Compared to psychology or biology, the replication issues of computational works have reasonable and efficient solutions. But making sure that these solutions are adopted will not be solved by articles such as this one. Just like in psychology and biology, we have to modify the incentive of the researchers, by adopting exigences, enforced domain-wide, of what constitute an acceptable scientific computational work.

>>>>

Replicability is a cornerstone of science.  If an experimental result cannot be
re-obtained by an independent party, it merely becomes, at best, an observation
that may inspire future research (\cite{Mesirov:2010,osc:2015}). Replication
issues have received increased attention in recent years, with a particular
media focus on medicine and psychology (\cite{Iqbal:2016}).  One could think
that computer science would mostly be shielded from such issues, since a
computer program intrinsically describes precisely what it does and is easy to
transmit perfectly to other researchers.

@@

A \emph{re-runnable} code is one that can be run again when needed, i.e. more than the one time that was needed to produce the results. It is important to notice that the re-runnability of a code is not an intrinsic property. Rather, it depends on the context, and re-runnability becomes increasingly difficult as the code ages. Therefore, to be re-runnable in other researchers' computers and through time, a re-runnable code should describe---with enough details to be recreated---an execution environment in which it is executable. As shown by \citep{Collberg:2016}, this is far from being either obvious or easy.\\

@@

Setting seeds is should be done carefully. Using 439 as a seed in the previous program would result in ten consecutive +1 steps\footnote{With CPython 3.3-3.6. See the next section for details.}, which---although a perfectly valid random walk---lend itself to a gross misinterpretation of the overall dynamic of the algorithm. Verifying that the qualitative aspects of the results and the conclusions that are made are not tied to a specific initialization of the pseudo-random generator is an integral part of any scientific undertaking in computational science;
this is usually done by repeating the simulations multiple times with different seeds.
%FIXME: is `simulations` pertinent here?

@@

% scripts to generates figures
The code should also contain all the scripts that were used to produce the figures. While it may be considered extraneous, code for plotting figures may contain non-trivial statistics code, and is actually just as vulnerable to bugs. With a code that is re-runnable, repeatable and contains such scripts, a user can quickly check that published figures can be reproduced.\\

@@

% reproducibility implies availability
It should be obvious by now that \emph{reproducibility implies availability}. As shown in \citep{Collberg:2016}, code is often unavailable, or only available upon request. While the latter may seem sufficient, changes in email address, changes in career, retirement, a busy inbox or poor archiving practices can make a code just as unreachable. Code \emph{and} input \emph{and} result files \emph{and} figure data should be available with the published article, as supplementary data, or through a DOI link to a scientific repository such as Figshare or Zenodo\footnote{Online code repositories such as GitHub \emph{are not} scientific repositories, and may disappear, change name or change their access policy at any moment. Direct links to them are not perpetual, and, when used, they should always be supplemented by a DOI link to a scientific archive.}.\\

@@

Reusability of course directly benefits other researchers from your team and outside of it. The easier it is to use your code, the lower the threshold is for other to study, modify and extend it. Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage. Moreover, scientists prefer to base their work on results they can verify. A reproducible and reusable code offers a platform both validated and easy-to-use, fostering the development of derivative works by other researchers. Those derivative works contribute to the impact of your original contribution.\\

@@

Despite all this, reusability is often overlooked, and it is not hard to see why. Scientists are rarely trained in software engineering, and reusability can represent an expensive endeavour if undertaken as an afterthought, for little tangible short-term benefits, for a codebase that might, after all, see only a single use. And, in fact, reusability is not as indispensable a requirement as the other ones presented here.\\

Yet, some simple measures can tremendously increase reusability, and at the same time strengthen reproducibility and re-runnability.\\

@@

Similarly, code behavior should be changed by commenting/uncommenting code \citep{Wilson:2017}. Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user. This improves reproducibility in two ways: it allows those conditions to be explicitly tracked in the result files since they contain the parameters used. It also allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper. With a documention explaining which script or configuration file correspond to which experiment, reproducing the different figures becomes straightforward.\\
%FIXME repeating what was said in reproducible.

Documentation, in fact, is one of the most potent tools for reusability. A proper documentation on how to install and run the software often makes the difference for other researchers whether they manage to use the code or not.\\

Furthermore, a comment describing what each function does, however evident, can avoid hours of head-scratching. Great code may need few comments. Scientists, however, are not always brilliant developers. Code quality, in science, averages on mediocre. Of course, bad, complicated code should be rewritten until is simple enough to explain itself. But realistically, this is hardly ever going to be done: there is simply not enough incentive for it. There, a comment that explains the intention and scientific reasons behind a block of code can tremendously useful.\\

@@

Re-usability is not an absolute requirement for scientific code. But it has many benefits, and a few simple measures can foster it considerably.

@@

Here we are able to replicate exactly the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\

But there are still more subtle problems with the description given above.
If we look more closely at it, we can realize
that nothing is said about the specific case of {\tt 0} when generating a step.
Do we have to consider {\tt 0} to be a positive or a negative step? Without
further information and without the original code, it is up to the reader to
decide. Likewise, the description is ambiguous regarding the first element of the walk. Is the initialization value included (it was not in our codes so far)? This slight difference changes the statistics, especially of short runs.\\

There is no better way to ferret out the ambiguities, big and small, of an article than to replicate it. This is one of the reasons why the \href{https://rescience.github.io}{ReScience
  journal} has been created. This journal {\em targets computational research and
encourages the explicit replication of already published research, promoting
new and open-source implementations in order to ensure that the original
research is reproducible}.\\

@@

%FIXME: can still be improved
Through a small random walk example using Python, we've seen a variety of
pitfalls that are easy to remain unaware of while writing scientific code. The
R$^0$ form, while easy to write and simple to understand is not a good
scientific code. Adding a single comment, we transformed this code into the
R$^1$ version (rerunnable) that partially solved the common dependencies
problem and will allow us to re-run the program two months or two years from
now. In the meantime, the R$^2$ version (repeatable) guarantees us (to some
extent) we can faithfully repeat our experiment from one run to the other.  Of
course, the full-fledge R$^4$ (reusable) form is much preferred but it requires
a fair amount of work that might be too much of a burden. R$^3$ (reproducible)
form seems to be an acceptable compromise and should be accepted as the minimum
scientific standard \citep{Wilson:2017}. This means this should be actually
checked by reviewers and publishers when code is part of a work worth being
published. But it's hardly the case today.\\

@@

Compared to psychology or biology, the replication issues of computational works have reasonable and efficient solutions. But making sure that these solutions are adopted will not be solved by articles such as this one. Just like in psychology and biology, we have to modify the incentive of the researchers, by adopting exigences, enforced domain-wide, of what constitutes an acceptable scientific computational work.
